---
title: "ESM 244 Lab 4"
author: "Tom Paschos"
date: "1/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Lab Week 4:

- Nonlinear least squares
- Panel regression level 0 example
- Shiny example with navbarPage

0. Attach Packages

```{r Load_Packages}

library(tidyverse)
library(Ecdat)
library(plm)
library(lmtest)
library(car)

```

1. Nonlinear least squares (logistic growth of bacteria)

```{r Load_Data}

CellGrowth <- read_csv("CellGrowth.csv")

# Look at it:

ggplot(CellGrowth, aes(x = Time, y = CellCount)) +
  geom_point()

# Find estimates for logistic gwoth:
# N =  A/(1+Be^(-rt))
# 3 parameters: A, B, and r

# Estimate for carrying capacity (K): 3700
# OK, if N0 is 2000, what is B? ~0.85
# r initial estimate: ~ 0.1035

BEstimate <- (3700 -2000)/2000

# Looks like there's exponential growth from Hour 0 - Hour 3.5 or 4
# Now, we're going to estimate r (the growth rate)

lag_phase <- CellGrowth %>% 
  filter(Time < 5)

ggplot(lag_phase, aes(x = Time, y = log(CellCount))) +
  geom_point()



lm(log(CellCount) ~ Time, data = lag_phase)
# coefficient for t = 0.1035; this is the r value

```

Find model A, B and r using Gauss-Newton method for iterative convergence (nonlinear least squares):

Use the nls() function: 

```{r nls}

# We can use anything for our model structure, making this function flexible.
# Trying to model the relationship of Cell Count and Time but we need A, B and r

cell_fit <- nls(CellCount ~ A/(1 + B*exp(-r*Time)),
                start = list(A = 3900, B = 0.85, r = 0.1035),
                data = CellGrowth,
                trace = TRUE)

# Gives you a traceback of all the iterations, values on the left column are the square residuals until they stop changing (convergence) at the bottom line
# The other columns are now your A, B and r values by running the model iteratively

cell_fit
# This shows us the values for those A, B and r values

A <- coef(cell_fit)[1] # This allows us to call A in case our data changes
B <- coef(cell_fit)[2]
r <- coef(cell_fit)[3]

# Create a new sequence of time values (time_seq)

time_seq <- seq(0,20,length = 100) # length = 100 makes 100 observations in the sequence

# Plug that new sequence into my model with the parameters A, B and r that we found:

cell_pred <- A/(1 + B*exp(-r*time_seq))

# Bind that together with the time_seq data:

pred_df <- data.frame(time_seq, cell_pred)

#Graph it with original data and our model predictions:
ggplot(CellGrowth, aes(x = Time, y = CellCount)) +
  geom_point(color = "blue", size = 3) +
  geom_line(data = pred_df, aes(x = time_seq, y = cell_pred), color = "orange", size = 1)


```

2. Panel Regression - First Shot

Variable we're really trying to understand is the average price

State: our entity variable
Year: our time variable
packpc: our dependent variable
avgprc: our predictor variable of interest

```{r Panel_Regression}

cigs_panel <- Cigarette %>% 
  select(state, year, packpc, avgprs)

# Let's examine the relationship b/w price and packs purchased

ggplot(cigs_panel, aes(x = avgprs, y = packpc, group = state)) +
  geom_point(aes(color = state)) +
  geom_smooth(method = "lm", aes(color = state), se = FALSE, size = 0.2)

ggplot(cigs_panel, aes(x = year, y = packpc, group = state)) +
  geom_point(aes(color = state)) +
  geom_smooth(method = "lm", aes(color = state), se = FALSE, size = 0.2)

```

Entity Fixed Effects model:

```{r EFE_model}

cigs_model <- plm(packpc ~ avgprs,
                  data = cigs_panel,
                  index = c("state","year"),
                  model = "within") # within says do an entity fixed effect model

cigs_model

# Interpretation:
# For every dollar increment in price of packaging we would expect a 0.32 less packages per person
# Per capita cigarette purchases will decrease by 0.32

coeftest(cigs_model, vcov. = vcovHC(cigs_model, type = "HC1"))

# SE term is often what people pull out of this.  The Estimate is slightly different.

```

Add time fixed effects:

```{r TFE_model}

et_cigs_model <- plm(packpc ~ avgprs,
                     data = cigs_panel,
                     index = c("state","year"),
                     model = "within",
                     effect = "twoways")

et_cigs_model
# coefficient has changed a little bit
# has including a TFE improved my model?  Let's use an pFtest to check:

pFtest(et_cigs_model, cigs_model) # order counts - more complete model first (time fixed effect model)

```




